---
title: "Dashbaord Q1 - Locations"
subtitle: "Labelled and interactive maps"
author: "Carlos Guío"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: journal
    highlight: tango
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(include = FALSE, echo = FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.showtext = TRUE, fig.retina = 1, dpi = 300, out.width = "90%")


```


## Read and prepare Q1 datasets

Q1 datasets are loaded by sourcing the `Prepare_Locations_Summaries.R`script. These datasets were designed to answer the questions:

* Who adopted/disadopted practices?
* Who accessed/used climate information services?
* Which years/ practices were surveyed at each site?

The Q1_practices and Q1_services datasets were processed to create summaries from them.


```{r source, echo = TRUE}
library(devtools)

source_url("https://raw.githubusercontent.com/cmguiob/CSA_Dashboard/main/R/Prepare_Locations_Summaries.R")

```

## Visualization

The following chunks have code to plot  located summaries about the survey (years surveyed, communities at each site and practices assessed ) and the main results of it, such as the main weather shocks experienced by households, and the percentages of adoption of practices -and use of CIS- by households and farmers. 


[Reference for world_data with ggplot](https://ggplot2.tidyverse.org/reference/map_data.html)

### Visualization choice: maps with `ggplot` (CRAN)

```{r function_map, echo = TRUE, include = TRUE}

f_locations <- function(dataset = locations_dat, regn, countr, cat){
  
  library(tidyverse)
  library(ggplot2)
  library(ggrepel)
  
  dataset2 <-  dataset %>%
        mutate(site = case_when(site == "Cinzana" ~ "Cinzana (MLI)",
                                site == "Fakara" ~ "Fakara (NER)",
                                site == "Kaffrine" ~ "Kaffrine (SEN)",
                                site == "Lawra-Jirapa" ~ "Lawra-Jirapa (GHA)",
                                site == "Bardiya" ~ "Bardiya (NPL)",
                                site == "Barisal" ~ "Barisal (BGD)",
                                site == "Khulna" ~ "Khulna (BGD)",
                                site == "Mahottari" ~ "Mahottari (NPL)",
                                site == "Nawalparasi" ~ "Nawalparasi (NPL)",
                                site == "Cauca" ~ "Cauca (COL)",
                                site == "El Tuma-La Dalia" ~ "El Tuma-La Dalia (NIC)",
                                site == "Santa Rita" ~ "Santa Rita (HND)",
                                site == "Olopa" ~ "Olopa (GTM)",
                                site == "Basona Werana" ~ "Basona Werana (ETH)",
                                site == "Doyogena" ~ "Doyogena (ETH)",
                                site == "Hoima" ~ "Hoima (UGA)",
                                TRUE ~ site)) %>%
        filter(region == regn,
               country == countr,
               category == cat)
        

  world_data <- ggplot2::map_data('world')%>%
        rename(country = region) %>%
        select(-subregion) %>%
        filter(country %in% dataset2$country)
  
  
  # Optional data to label countries direct over their map
  world_labels <- world_data %>%
        mutate(country = case_when(country == "Bangladesh" ~ "BGD",
                                   country == "Colombia" ~ "COL",
                                   country == "Ethiopia" ~ "ETH",
                                   country == "Ghana" ~ "GHA",
                                   country == "Guatemala" ~ "GTM",
                                   country == "Honduras" ~ "HND",
                                   country == "Mali" ~ "MLI",
                                   country == "Nepal" ~ "NPL",
                                   country == "Nicaragua" ~ "NIC",
                                   country == "Niger" ~ "NER",
                                   country == "Senegal" ~ "SEN",
                                   country == "Uganda" ~ "UGA",
                                   TRUE ~ country)) %>%
        group_by(country) %>%
        summarise(long = 
                  case_when(country == "BGD" ~ mean(long),
                            country == "COL" ~ mean(long),
                            country == "ETH" ~ mean(long),
                            country == "GHA" ~ mean(long),
                            country == "GTM" ~ mean(long),
                            country == "HND" ~ mean(long),
                            country == "MLI" ~ mean(long),
                            country == "NPL" ~ mean(long),
                            country == "NIC" ~ mean(long),
                            country == "NER" ~ mean(long),
                            country == "SEN" ~ mean(long),
                            country == "UGA" ~ mean(long)), 
                 lat = 
                 case_when(country == "BGD" ~ mean(lat),
                           country == "COL" ~ mean(lat),
                           country == "ETH" ~ mean(lat),
                           country == "GHA" ~ mean(lat),
                           country == "GTM" ~ mean(lat),
                           country == "HND" ~ mean(lat),
                           country == "MLI" ~ mean(lat),
                           country == "NPL" ~ mean(lat),
                           country == "NIC" ~ mean(lat),
                           country == "NER" ~ mean(lat),
                           country == "SEN" ~ mean(lat),
                           country == "UGA" ~ mean(lat)))
  # Plot
  ggplot(data = world_data,
         aes(x = long, 
             y = lat)) +
    geom_polygon(data = world_data,
                 aes(group = group),
                 color = "#faf9f5", 
                 fill = "#ECEBE7", 
                 size = 1) +
    #geom_text(data = world_labels, 
    #          aes(label = country),
    #          color = "#B8B7B3",
    #          size = 8,
    #          face = "bold")+
    geom_point(data = dataset2,
               aes(x = longitude_site, 
                   y = latitude_site, 
                   size = n,
                   color = site),
               shape = 19,
               #size = 6,
               alpha = 0.5)+
      scale_size_continuous(range = c(7,10))+
      scale_color_hue(l = 70, c = 40, h = c(45, 250))+
      scale_fill_hue(l = 70, c = 40, h = c(45, 250))+
      geom_text_repel(data = dataset2,
                     aes(label = 
                           if(category == "Years"){
                             paste0("Years surveyed:\n", details)
                             } 
                         else if(category == "Communities"){
                           paste(n, "communities surveyed")
                           }
                         else if(category == "Practices"){
                           paste("Practices assessed:\n",details)
                           }
                         else if(category == "Households"){
                           paste(n,"households surveyed.\n",details)
                         }
                         else if(category == "Farmers"){
                           paste(n,"farmers surveyed.\n",details)
                         }
                         else {paste0(n, "% of households had shocks.\n From them:\n",details)},
                         x = longitude_site, 
                         y = latitude_site,
                         color = site),
                     show_guide=F,
                     size = 3,
                     #fontface = "bold",
                     box.padding = if (dataset2$region == "South Asia"){2.5} else {3.5},
                     force = if (dataset2$region == "Latina America"){3.5} 
                             else if (dataset2$region == "South Asia") {4}
                             else if (dataset2$region == "West Africa") {3}
                             else {4},
                     point.padding = 1.5,
                     #nudge_y = -0.5,
                     nudge_x = 0.1,
                     #color = "#5D5C58",
                     max.overlaps = Inf,
                     direction = "both",
                     xlim = c(-Inf, Inf))+
    guides(size = "none",
          color = guide_legend(override.aes = list(size = 8)))+
    labs(color = "Sites",
         caption = if(dataset2$category == "Practices"){"*m = practice has mitigation features, g: gender impact"})+
    theme_minimal()+
    theme(panel.grid = element_blank(),
          # Optional to use same background as in dashboard
          #plot.background = element_rect(fill = "#faf9f5", color = "#faf9f5"),
          legend.text = element_text(size = 11, 
                                      color = "#5D5C58"),
          legend.title = element_text(size = 11,
                                      color = "#B8B7B3",
                                      face = "bold"),
          legend.key.size = unit(1.5, 'lines'),
          legend.justification = "top",
          legend.position = "bottom",
          plot.caption = element_text(size = 7, 
                                      hjust = 1, 
                                      margin = margin(5,0,5,0, "mm")),
          axis.text = element_blank(),
          axis.title = element_blank(),
          plot.margin = unit(if (dataset2$region == "South Asia"){c(0,20,0,-20)}
                             else if (dataset2$region == "West Africa"){c(0,15,0,35)}
                             else {c(0,10,0,15)}, "mm"))+
      coord_quickmap(clip = "off")
  
  
}


```


### Use the function to create a map with summaries

The argumens of the function take in values similar to the drop down menus used in the PBI-dashboard. The argument `rgn`: region, `countr`:country, `cat`:category.


```{r plot_ggsankey, include= TRUE, echo = TRUE}

p_locations <- f_locations(regn = "East Africa", countr = "Ethiopia", cat = "Practices")

p_locations

# Export
#ggsave(file = "p_sankey.png", plot = p_sankey, device = "png", type = "cairo", path = "Input a path from your PC, dpi = 300, width = 9.5, height = 5)

```
