---
title: "Dashboard: Effects"
subtitle: "Tornado plots"
author: "Carlos Guío"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: journal
    highlight: tango
editor_options: 
  chunk_output_type: inline
---



```{r setup, include=FALSE}

knitr::opts_chunk$set(include = FALSE, echo = FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.showtext = TRUE, fig.retina = 1, dpi = 300, out.width = "90%")
```


## Read and prepare Q1 and Q2 datasets

Q1 and Q2 datasets are loaded by sourcing the `Prepare_Effects_Tornado.R`script. 

Q1 datasets were designed to answer questions such as:

* Who adopted/disadopted practices and why?
* Who accessed/used climate information services?

Q2 datasets were designed to answer questions such as:

* How does the implementation of CSA practices affect food security, vulnerability to climate shocks, gender and income?

* How does the use of CIS affect changes in animal and crop management?

The Q1 and Q2 datasets were lightly processed to integrate them in a single dataset, which can be sourced with the following chunk:


```{r source, echo = TRUE}
library(devtools)

source_url("https://raw.githubusercontent.com/cmguiob/CSA_Dashboard/main/R/Prepare_Effects_Tornado.R")

```

### Visaulization choice: tornado plot with `ggplot` (CRAN)

In the next chunk you will find a function that allows to plot the relationship between practices (or services) and perceived socio-economic effects (indicators) at each site. The information is at the farmer level.

```{r plot_tornado, include = TRUE}

f_tornado <- function(dataset = effects_dat, CSA_category, outcome, indicator, location, year_1, year_2){
  
  library(tidyverse)
  library(ggplot2)
  
   # Error messages
  if (!CSA_category %in% c("Practices", "Services")) 
    stop("use either Practices or Services as CSA_category")

  if (!outcome %in% dataset$outcome_category) 
    stop("assign a valid outcome category")

  if (!indicator %in% dataset$indicators) 
    stop("assign a valid indicator")
  
  years <- dataset %>% ungroup() %>%
    filter(site %in% location, option == CSA_category, outcome_category == outcome, 
         indicators == indicator) %>%            
    select(year) %>%
  distinct()

  if (!any(range(years) %in% seq(year_1, year_2, by=1))) 
    stop("assign a valid year interval to the surveyed site")

  if (!location %in% dataset$site) 
    stop("assign a valid site name as location")
  
  
  tornado_dat <- dataset %>%
         filter(outcome_category == outcome,
                indicators == indicator,
                site == location)
         filter(!is.na(practice_type)) %>%
         mutate(percentage_lab = case_when(percentage < 2 ~ "",
                                       TRUE ~ as.character(paste0(percentage, "%"))),
                gender_legend = paste0(gender," (n: ",n_site,")"))
  
}

tornado_dat <- effect_dat %>%
         mutate(percentage = case_when(percentage < 10 ~ "",
                                       TRUE ~ as.character(paste0(percentage, "%")))) %>%
         # This is a slicer in PBI
         filter(outcome_category == "Food security" & indicator == "Improved food access") %>%
         filter(site == "Doyogena")

ggplot(tornado_dat) +
  geom_col(aes(x = practice_type, y = ifelse(gender == "Male", n, -n), fill = gender, alpha = alpha)) +
  geom_text(aes(x = practice_type, y = ifelse(gender == "Male", n, -n), label = percentage),
            nudge_y = ifelse(tornado_dat$gender == "Male", 10, -10),
            size = 2) +
  scale_alpha(range = c(0.5, 1))+
  scale_fill_manual(values = c("#32aa52", "#a9713d")) +
  scale_y_continuous(limits = c(-max(tornado_dat$n) -10, max(tornado_dat$n) +10))+
  facet_grid(indicator ~ site + year)+
  coord_flip() +
  guides(alpha = "none")+
  theme_minimal()+
  theme(panel.grid = element_blank(),
        panel.background = element_rect(fill = "#faf9f5", color = "#ECEBE7" ),
        plot.background = element_rect(fill = "#faf9f5", 
                                       color = "#faf9f5"),
        strip.background = element_rect(fill = "#ECEBE7", color = "#ECEBE7"),
        strip.text = element_text(margin = margin(2.5,0,2.5,0, "mm"), size = 8, color = "#5D5C58"),
        legend.title = element_blank(),
        legend.justification = "left",
        legend.position = "bottom",
        legend.text = element_text(size = 8, color = "#5D5C58"),
        axis.title = element_blank(),
        axis.text.y = element_text(color = "#B8B7B3", size = 7),
        axis.text.x = element_blank())

```


