---
title: "Dashboard effects"
subtitle: "Tornado plots"
author: "Carlos Guío"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: journal
    highlight: tango
editor_options: 
  chunk_output_type: inline
---

The goal of this plot is to show which the relationship between practices and perceived social and environmental effects at each site. The information is at the farmer level.


```{r setup, include=FALSE}

knitr::opts_chunk$set(include = FALSE, echo = FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.showtext = TRUE, fig.retina = 1, dpi = 300, out.width = "75%")
```


```{r load}
library(tidyverse)


Q1_practices <- read_csv(here::here("Data","EU-IFAD","Q1_practices.csv")) %>%
                select(location_type, 
                       location_name, 
                       farmer_id, 
                       farmer_gender,
                       year, 
                       practice_type,
                       subtype,
                       adopted_cases) %>%
                pivot_wider(names_from = location_type, values_from = location_name) %>%
                unchop(everything()) %>%
                mutate(option = "Practices",
                       farmer_id = case_when(
               farmer_id == "337a45e573817fba21d251ecc3a37523e14cc62d073" & 
               farmer_gender == "Female" ~ "337a45e573817fba21d251ecc3a37523e14cc62d073_F",
               farmer_id == "3e2a4c557a0a5277638e42df23b17bd63eb41858faa" & 
               farmer_gender == "Female" ~ "3e2a4c557a0a5277638e42df23b17bd63eb41858faa_F",
               farmer_id == "4e6a1f557611aec6b347a05cde3a1e093fb88e883a0" & 
               farmer_gender == "Female" ~ "4e6a1f557611aec6b347a05cde3a1e093fb88e883a0_F",
               farmer_id == "533a330570d0a9e20fdd55e40993645a7c9838e11b0" & 
               farmer_gender == "Female" ~ "533a330570d0a9e20fdd55e40993645a7c9838e11b0_F",
               farmer_id == "27fa75957420473f9ad14deb5ed9b7701031f701149" & 
               farmer_gender == "Female" ~ "27fa75957420473f9ad14deb5ed9b7701031f701149_F",
               farmer_id == "a75acb457efe9fb18b082f3a52d2dad46a69a42cf8e" & 
               farmer_gender == "Female" ~ "a75acb457efe9fb18b082f3a52d2dad46a69a42cf8e_F",
               farmer_id == "112a7c157220d8de79f29b9ac7bf6af20daa0a05983" & 
               farmer_gender == "Female" ~ "112a7c157220d8de79f29b9ac7bf6af20daa0a05983_F",
               farmer_id == "fd7a65f575e4ac103cbe3b6832d559cbc199d82101b" & 
               farmer_gender == "Female" ~ "fd7a65f575e4ac103cbe3b6832d559cbc199d82101b_F",
               TRUE ~ farmer_id),
               practice_type = str_to_sentence(practice_type))

Q1_services <- read_csv(here::here("Data","EU-IFAD","Q1_services.csv")) %>%
              select(location_type, 
                       location_name, 
                       farmer_id, 
                       farmer_gender,
                       year, 
                       practice_type,
                       subtype,
                       adopted_cases) %>%
                pivot_wider(names_from = location_type, values_from = location_name) %>%
                unchop(everything()) %>%
                mutate(option = "Services",
                       farmer_id = case_when(
               farmer_id == "337a45e573817fba21d251ecc3a37523e14cc62d073" & 
               farmer_gender == "Female" ~ "337a45e573817fba21d251ecc3a37523e14cc62d073_F",
               farmer_id == "3e2a4c557a0a5277638e42df23b17bd63eb41858faa" & 
               farmer_gender == "Female" ~ "3e2a4c557a0a5277638e42df23b17bd63eb41858faa_F",
               farmer_id == "4e6a1f557611aec6b347a05cde3a1e093fb88e883a0" & 
               farmer_gender == "Female" ~ "4e6a1f557611aec6b347a05cde3a1e093fb88e883a0_F",
               farmer_id == "533a330570d0a9e20fdd55e40993645a7c9838e11b0" & 
               farmer_gender == "Female" ~ "533a330570d0a9e20fdd55e40993645a7c9838e11b0_F",
               farmer_id == "27fa75957420473f9ad14deb5ed9b7701031f701149" & 
               farmer_gender == "Female" ~ "27fa75957420473f9ad14deb5ed9b7701031f701149_F",
               farmer_id == "a75acb457efe9fb18b082f3a52d2dad46a69a42cf8e" & 
               farmer_gender == "Female" ~ "a75acb457efe9fb18b082f3a52d2dad46a69a42cf8e_F",
               farmer_id == "112a7c157220d8de79f29b9ac7bf6af20daa0a05983" & 
               farmer_gender == "Female" ~ "112a7c157220d8de79f29b9ac7bf6af20daa0a05983_F",
               farmer_id == "fd7a65f575e4ac103cbe3b6832d559cbc199d82101b" & 
               farmer_gender == "Female" ~ "fd7a65f575e4ac103cbe3b6832d559cbc199d82101b_F",
               TRUE ~ farmer_id),
               practice_type = str_to_sentence(practice_type)) 

Q2_practices <- read_csv(here::here("Data","EU-IFAD","Q2_practices.csv")) %>%
  select(year,
         site,
         farmer_id,
         farmer_gender,
         practice_type,
         outcome_category,
         indicator,
         answer) %>%
  mutate(option = "Practices",
         practice_type = str_to_sentence(practice_type),
         indicator = case_when(
                    indicator == "CSA option increased production" ~ "Increased production",
                    indicator == "CSA driven decrease in vulnerability" ~ 
                      "Reduced climate vulner.",
                    indicator == "Decreased agricultural labour time" ~ "Decreased labour time",
                    indicator == "Increased agricultural labour time" ~ "Increased labour time",
                    indicator == "Participation on use of CSA income" ~ 
                      "Participated on CSA-income use",
                    TRUE ~ indicator)) 

Q2_services <- read_csv(here::here("Data","EU-IFAD","Q2_services.csv")) %>%
  select(year,
         site,
         farmer_id,
         farmer_gender,
         practice_type = service_type,
         outcome_category,
         indicator,
         answer) %>%
  mutate(option = "Services",
         practice_type = str_to_sentence(practice_type))


```

Prepare dat includes counting the total number of farmers at each site, disaggregated by gender, calculating the number of farmers at each site who answered yes for each combination of practice and indicator, the percentages of these farmers and, using the percentages, the top 3 practices that were most important for each indicator. Percentages are calculated within the group of people that adopted practices.

```{r prep}

# Number of farmers by gender that implemented practices at each site
n_site_practices <- Q1_practices %>%
  filter(!is.na(farmer_gender)) %>% 
  filter(adopted_cases == "Implemented") %>% 
  select(site, farmer_gender, farmer_id) %>% 
  group_by(site, farmer_gender) %>%
  distinct(farmer_id) %>%
  summarise(n_site = n()) %>%
  ungroup()

percent_practices <- Q2_practices %>%
  filter(answer == "yes")  %>%
  select(year, site, farmer_id, farmer_gender, outcome_category, practice_type, indicator) %>%
  group_by(year, 
           site, 
           outcome_category, 
           practice_type, 
           indicator, 
           farmer_gender) %>%
  distinct() %>% 
  count() %>%
  ungroup() %>%
  left_join(n_site_practices, by = c("site", "farmer_gender")) %>%
  mutate(percentage = round(n*100/n_site,1))
  

top3_practices <- percent_practices %>%
  group_by(year, site, outcome_category, farmer_gender, indicator) %>%
  top_n(3, wt = percentage) %>%
  mutate(alpha = "*")  %>%
  select(-n, -n_site, -percentage) %>%
  ungroup()

effect_practices <- left_join(percent_practices, 
                        top3_practices, 
                        by = c("year","site","outcome_category", "practice_type", "indicator", "farmer_gender")) %>%
              #Alpha for the ggplot
              mutate(alpha = case_when(alpha == "*" ~ 1,
                                       TRUE ~ 0.8))
effect_practices

```


```{r}

# Number of farmers by gender that accessed and used services at each site
n_site_services <-Q1_services %>%
  filter(!is.na(farmer_gender)) %>% 
  filter(adopted_cases == "Accessed and used it") %>% 
  select(site, farmer_gender, farmer_id) %>% 
  group_by(site, farmer_gender) %>%
  distinct(farmer_id) %>%
  summarise(n_site = n()) %>%
  ungroup()


percent_services <- Q2_services %>%
  filter(answer == "yes")  %>%
  select(year, site, farmer_id, farmer_gender, outcome_category, practice_type, indicator) %>%
  group_by(year, 
           site, 
           outcome_category, 
           practice_type, 
           indicator, 
           farmer_gender) %>%
  distinct() %>% 
  count() %>%
  ungroup() %>%
  left_join(n_site_services, by = c("site", "farmer_gender")) %>%
  mutate(percentage = round(n*100/n_site,1))
  

top3_services <- percent_services %>%
  group_by(year, site, outcome_category, farmer_gender, indicator) %>%
  top_n(3, wt = percentage) %>%
  mutate(alpha = "*")  %>%
  select(-n, -n_site, -percentage) %>%
  ungroup()

effect_services <- left_join(percent_services, 
                        top3_services, 
                        by = c("year","site","outcome_category", "practice_type", "indicator", "farmer_gender")) %>%
              #Alpha for the ggplot
              mutate(alpha = case_when(alpha == "*" ~ 1,
                                       TRUE ~ 0.8))
effect_services
```




```{r save}

effect_dat <- bind_rows(effect_practices %>% mutate(option = "Practices"),
                        effect_services %>% mutate(option = "Services") )

write_csv(effect_dat, here::here("Data","EU-IFAD","effects.csv"))

```

## Including Plots



```{r plot_tornado, include = TRUE}

tornado_dat <- effect_dat %>%
         mutate(percentage = case_when(percentage < 10 ~ "",
                                       TRUE ~ as.character(paste0(percentage, "%")))) %>%
         # This is a slicer in PBI
         filter(outcome_category == "Food security" & indicator == "Improved food access") %>%
         filter(site == "Doyogena")

ggplot(tornado_dat) +
  geom_col(aes(x = practice_type, y = ifelse(farmer_gender == "Male", n, -n), fill = farmer_gender, alpha = alpha)) +
  geom_text(aes(x = practice_type, y = ifelse(farmer_gender == "Male", n, -n), label = percentage),
            nudge_y = ifelse(tornado_dat$farmer_gender == "Male", 10, -10),
            size = 2) +
  scale_alpha(range = c(0.5, 1))+
  scale_fill_manual(values = c("#32aa52", "#a9713d")) +
  scale_y_continuous(limits = c(-max(tornado_dat$n) -10, max(tornado_dat$n) +10))+
  facet_grid(indicator ~ site + year)+
  coord_flip() +
  guides(alpha = "none")+
  theme_minimal()+
  theme(panel.grid = element_blank(),
        panel.background = element_rect(fill = "#faf9f5", color = "#ECEBE7" ),
        plot.background = element_rect(fill = "#faf9f5", 
                                       color = "#faf9f5"),
        strip.background = element_rect(fill = "#ECEBE7", color = "#ECEBE7"),
        strip.text = element_text(margin = margin(2.5,0,2.5,0, "mm"), size = 8, color = "#5D5C58"),
        legend.title = element_blank(),
        legend.justification = "left",
        legend.position = "bottom",
        legend.text = element_text(size = 8, color = "#5D5C58"),
        axis.title = element_blank(),
        axis.text.y = element_text(color = "#B8B7B3", size = 7),
        axis.text.x = element_blank())

```


