---
title: "Dashbaord Uptake"
subtitle: "Treemap/mosaic plots"
author: "Carlos Guío"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: journal
    highlight: tango
editor_options: 
  chunk_output_type: inline
---

The goal of this plot is to show which practices are available at a given site and the percentage of this practices that were adopted, disadopted and not adopted by male/female headed households.


```{r setup}
knitr::opts_chunk$set(include = FALSE, echo = FALSE, warning = FALSE, message = FALSE, fig.align="center", fig.showtext = TRUE, fig.retina = 1, dpi = 300, out.width = "75%")
```



```{r load, echo=FALSE}

library(tidyverse)

#Read
Q1_practices_w <- read_csv(here::here("Data","EU-IFAD","Q1_practices.csv")) %>%
  select(year,
         location_type, 
         location_name, 
         farmer_id,
         gender = farmer_gender,
         practice_type,
         subtype,
         adopted_cases,
         description) %>%
  pivot_wider(names_from = location_type, values_from = location_name) %>%
  #unnest columnas que quedaron anidadas
  unchop(everything()) %>%
  select(-community) %>%
  distinct() %>%
  mutate(option = "Practices",
         subtype = case_when(
                       subtype == "Integrated nutrient management" ~ "Integrated management",
                       subtype == "Terraces biological measures" ~ "Terraces (bio-measures)",
                       TRUE ~ subtype),
        farmer_id = case_when(
               farmer_id == "337a45e573817fba21d251ecc3a37523e14cc62d073" & 
               gender == "Female" ~ "337a45e573817fba21d251ecc3a37523e14cc62d073_F",
               farmer_id == "3e2a4c557a0a5277638e42df23b17bd63eb41858faa" & 
               gender == "Female" ~ "3e2a4c557a0a5277638e42df23b17bd63eb41858faa_F",
               farmer_id == "4e6a1f557611aec6b347a05cde3a1e093fb88e883a0" & 
               gender == "Female" ~ "4e6a1f557611aec6b347a05cde3a1e093fb88e883a0_F",
               farmer_id == "533a330570d0a9e20fdd55e40993645a7c9838e11b0" & 
               gender == "Female" ~ "533a330570d0a9e20fdd55e40993645a7c9838e11b0_F",
               farmer_id == "27fa75957420473f9ad14deb5ed9b7701031f701149" & 
               gender == "Female" ~ "27fa75957420473f9ad14deb5ed9b7701031f701149_F",
               farmer_id == "a75acb457efe9fb18b082f3a52d2dad46a69a42cf8e" & 
               gender == "Female" ~ "a75acb457efe9fb18b082f3a52d2dad46a69a42cf8e_F",
               farmer_id == "112a7c157220d8de79f29b9ac7bf6af20daa0a05983" & 
               gender == "Female" ~ "112a7c157220d8de79f29b9ac7bf6af20daa0a05983_F",
               farmer_id == "fd7a65f575e4ac103cbe3b6832d559cbc199d82101b" & 
               gender == "Female" ~ "fd7a65f575e4ac103cbe3b6832d559cbc199d82101b_F",
               TRUE ~ farmer_id),
         practice_type = str_to_sentence(practice_type),
         subtype =str_to_sentence(subtype))


  # Code to clean names
  #%>%
  # Replace anything after quotes
  #mutate(practice = str_replace(practice, "(?<=[\"]).+", "")) %>%
  # Replace quotes
  #mutate(practice = str_replace(practice, "[\"]", "")) %>%
  # Format to sentence with first letter uppercase
  #mutate(practice = str_to_sentence(practice))%>%
  # Remove period at the end
  #mutate(practice = str_replace(practice, "[\\.]", "")) %>%
  # Remove space at the end
  #mutate(practice = str_trim(practice)) %>%
  #Replace anything after open parentheses
  #mutate(practice_gen = str_replace(practice, "(?<=\\().+", "")) %>%
  #Replace parentheses
  #mutate(practice_gen = str_replace(practice_gen, "\\(", "")) %>%
  # Replace anything after colon
  #mutate(practice_gen = str_replace(practice_gen, "(?<=\\:).+", "")) %>%
  #Replace colon
  #mutate(practice_gen = str_replace(practice_gen, "\\:", "")) %>%
  # Remove text after dash and space
  #mutate(practice_gen = str_replace(practice_gen, "(?<=[\\-])\\s.+", "")) %>%
  #Remove dash at the end of sentence
  #mutate(practice_gen = str_replace(practice_gen, "[\\-]$", "")) %>%
  #Remove space at the end
  #mutate(practice_gen = str_trim(practice_gen))

Q1_services_w <- read_csv(here::here("Data","EU-IFAD","Q1_services.csv")) %>%
  select(year,
         location_type, 
         location_name, 
         farmer_id,
         gender = farmer_gender,
         practice_type,
         subtype,
         adopted_cases,
         description) %>%
  filter(practice_type != "No service") %>%
  pivot_wider(names_from = location_type, values_from = location_name) %>%
  #unnest columnas que quedaron anidadas
  unchop(everything()) %>%
  select(-community) %>%
  distinct() %>%
  mutate(option = "Services",
         description = case_when(is.na(description)~ "No description available",
                                 TRUE ~ description),
         farmer_id = case_when(
               farmer_id == "337a45e573817fba21d251ecc3a37523e14cc62d073" & 
               gender == "Female" ~ "337a45e573817fba21d251ecc3a37523e14cc62d073_F",
               farmer_id == "3e2a4c557a0a5277638e42df23b17bd63eb41858faa" & 
               gender == "Female" ~ "3e2a4c557a0a5277638e42df23b17bd63eb41858faa_F",
               farmer_id == "4e6a1f557611aec6b347a05cde3a1e093fb88e883a0" & 
               gender == "Female" ~ "4e6a1f557611aec6b347a05cde3a1e093fb88e883a0_F",
               farmer_id == "533a330570d0a9e20fdd55e40993645a7c9838e11b0" & 
               gender == "Female" ~ "533a330570d0a9e20fdd55e40993645a7c9838e11b0_F",
               farmer_id == "27fa75957420473f9ad14deb5ed9b7701031f701149" & 
               gender == "Female" ~ "27fa75957420473f9ad14deb5ed9b7701031f701149_F",
               farmer_id == "a75acb457efe9fb18b082f3a52d2dad46a69a42cf8e" & 
               gender == "Female" ~ "a75acb457efe9fb18b082f3a52d2dad46a69a42cf8e_F",
               farmer_id == "112a7c157220d8de79f29b9ac7bf6af20daa0a05983" & 
               gender == "Female" ~ "112a7c157220d8de79f29b9ac7bf6af20daa0a05983_F",
               farmer_id == "fd7a65f575e4ac103cbe3b6832d559cbc199d82101b" & 
               gender == "Female" ~ "fd7a65f575e4ac103cbe3b6832d559cbc199d82101b_F",
               TRUE ~ farmer_id),
         practice_type = str_to_sentence(practice_type),
         subtype = str_to_sentence(subtype))

Q1_options_w <- bind_rows(Q1_practices_w, Q1_services_w)

coordinates <- read_csv(here::here("Data","EU-IFAD","CCAFS_Villages _coordinates_GPS.csv"))


```

```{r prep, include = TRUE, echo = TRUE}




```

Export table for use in PBI

```{r save}

write_csv(uptake_dat, here::here("Data","EU-IFAD","uptake.csv"))

```


According to Claus Wilke, humans are better perceiving distances over areas. Therefore, a bar chart is decoded easier than 

```{r barplot, include = TRUE, echo = TRUE}

library(tidyverse)
library(ggplot2)
library(ggrepel)

sankey_dat <- uptake_dat %>% 
  filter(subtype %in% c("Tree planting")) %>%
  filter(n != 0)%>%
  mutate(gender_n = paste(substr(gender,1,1), paste0("(",n_site,")"), sep = "\n"))%>%
  arrange(year, site, practice_type, subtype, gender, desc(adopted_cases)) %>%
  mutate(pos = cumsum(n) - (0.5 * n)) %>%
  ungroup() %>%
  group_by(site) %>%
  mutate(x = case_when(any(str_detect(gender, "Female-headed"))& gender == "Female-headed"~ 0.5,
                       any(str_detect(gender, "Female-headed"))& gender == "Male-headed"~ 1.5,
                       any(!str_detect(gender, "Female-headed"))& gender == "Male-headed"~ 0.5),
         xend = case_when(any(str_detect(gender, "Female-headed"))& gender == "Female-headed"~ 1.5,
                       any(str_detect(gender, "Female-headed"))& gender == "Male-headed"~ 2.5,
                       any(!str_detect(gender, "Female-headed"))& gender == "Male-headed"~ 1.5))
         
  

  ggplot(sankey_dat, aes(fill = adopted_cases, y = n, x = gender_n)) + 
  geom_bar(position="stack", stat="identity")+
  geom_label_repel(aes(label = paste(sprintf("%.1f", percentage),"%"), 
                x = gender_n, 
                y = pos), 
            size = 3.5,
            alpha = 0.6,
            force = 0.1,
            box.padding = 0.1,
            direction = "x",
            max.overlaps = Inf,
            fill = "white")+
  facet_grid(year ~ site, scales = "free_x", space = "free_x") +
  geom_segment(data = sankey_dat, 
               aes(x = x, 
                   xend = xend, 
                   y = n_site, 
                   yend = n_site), 
               linetype=2, color = "#B8B7B3")+
  scale_y_continuous(labels = scales::percent_format()) +
  #scale_fill_OkabeIto(order = c(3,4, 1,2))+
  guides(fill = guide_legend(override.aes = list(size = 7)))+
  theme_minimal()+
  theme(panel.grid = element_blank(),
        panel.background = element_rect(fill = "#faf9f5", color = "#ECEBE7" ),
        plot.background = element_rect(fill = "#faf9f5", 
                                       color = "#faf9f5"),
        strip.background = element_rect(fill = "#ECEBE7", color = "#ECEBE7"),
        legend.title = element_blank(),
        legend.justification = "top",
        legend.position = "left",
        axis.title = element_blank(),
        axis.text.y = element_blank()) 

```

Join with map

```{r map, include = TRUE, echo = TRUE}

library(ggplot2)
library(tidyverse)
library(ggrepel)

world_data <- ggplot2::map_data('world')%>%
  rename(country = region) %>%
  select(-subregion) %>%
  filter(country %in% uptake_dat$country)

ggplot(data = world_data,
       aes(x = long, 
           y = lat)) +
  geom_polygon(data = world_data,
               aes(group = group),
               color = "#faf9f5", 
               fill = "#ECEBE7", 
               size = 1) +
  geom_point(data = uptake_dat %>% 
               filter(subtype %in% c("Daily and sesonal")), 
             aes(x = longitude_site, 
                 y = latitude_site, 
                 color = subtype),
             shape = 19,
             alpha = 0.8)+
  geom_label_repel(data = uptake_dat %>% 
               ungroup() %>%
               filter(subtype %in% c("Daily and sesonal")) %>%
               distinct(site, .keep_all = TRUE),
                   aes(label = site,
                       x = longitude_site, 
                       y = latitude_site),
                   size = 4,
                   box.padding = 1,
                   point.padding = 0,
                   nudge_y = 0,
                   max.overlaps = Inf,
                   direction = "both",
                   xlim = c(-Inf, Inf),
                   alpha = 0.5)+
guides(color = "none")+
  theme_minimal()+
  theme(panel.grid = element_blank(),
        plot.background = element_rect(fill = "#faf9f5", 
                                       color = "#faf9f5"),
        axis.text = element_blank(),
        axis.title = element_blank())+
    coord_quickmap()


```

Descriptions in reactive text boxes

```{r text, include = TRUE, echo = TRUE}

library(tidyverse)
library(ggplot2)
library(ggtext)

uptake_dat %>%
filter(subtype %in% c("Agroforestry fallow")) %>%
 ggplot() +
  geom_textbox(
    aes(x = 0.1, y = 1, label = description),
    width = grid::unit(0.75, "npc"), # 73% of plot panel width
    hjust = 0, 
    vjust = 1,
    size = 7,
    fill = "#ECEBE7",
    colour = "#5D5C58",
    box.colour = "#5D5C58",
  ) +
  xlim(0, 1) + ylim(0, 1) +
  theme_minimal()+
  theme(panel.grid = element_blank(),
        plot.background = element_rect(fill = "#faf9f5", 
                                       color = "#faf9f5"),
        axis.title = element_blank(),
        axis.text = element_blank()) 


```

